---
title: "Week 2 Assignment - Data Cleaning"
author: "Liane Chen"
format: html
---
```{r, message=FALSE}
library(tidyverse)
library(lubridate)
library(RColorBrewer)
```

```{r}
datadir_processed <- "data/processed/"

snowcover_data <- read_csv(file.path(datadir_processed, snowcover_csv))

snowcover_data_fixed <- snowcover_data %>%
  mutate(snow_days = ifelse(Snow_cover > 10, 1, 0),
         Date2 = as_date(Date))
```

```{r}
snowcover_data_fixed <- snowcover_data_fixed %>%
  mutate(Date2 = ifelse(is.na(Date2),  dmy("08/06/06"), Date2))

snowcover_data_fixed
```
```{r}
snowcover_data_fixed <- snowcover_data %>%
  mutate(Date = ifelse(Date == "8&9 june 06", "8 june 06", Date),
         Date2 = dmy(Date))
```


1.  Clean the `Water_cover` column to transform it into the correct data type and respect expectations for a percentage
```{r}
snowcover_data_fixed %>%
  count(Water_cover) %>%
  filter(is.na(as.numeric(Water_cover)))
```

```{r}
snowcover_data_fixed %>%
  filter(Water_cover ==".") %>%
  View()
```

```{r}
snowcover_data_fixed <- snowcover_data_fixed %>%
  mutate(Water_cover = ifelse(Water_cover==".", NA, Water_cover))
```


2.  Clean the `Land_cover` column to transform it into the correct data type and respect expectations for a percentage
If 2% or less of your data is weird, getting rid of it is fine. Anything more, something is off. Just document it. 
Get rid of the lines with negative percentages. For lines with NA Snow cover and -100 Land Cover, that doesn't look right either. 

```{r}
snowcover_data_fixed %>%
  count(Land_cover) %>%
  filter(is.na(as.numeric(Land_cover)))
```

```{r}
snowcover_data_fixed %>%
  filter(Land_cover ==".") %>%
  View()
```

```{r}
snowcover_data_fixed <- snowcover_data_fixed %>%
  mutate(Land_cover = ifelse(Land_cover==".", NA, Land_cover))
```


3.  Use the relationship between the three cover columns (Snow, Water, Land) to infer missing values where possible and recompute the `Total_cover` column
